---
title: 'Individual Project: Red Wine Analysis'
author: "Chelsea Alford"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Loading Necessary Packages
```{r}
library(ggplot2) 
library(ggcorrplot) 
library(randomForest)
library(rpart) 
library(tree)
```



##DATA PREPARATION
```{r}
data <- read.csv('winequality-red.csv')
str(data)

#creating a data frame for visualizations with rounded values that are easier to display
datavis <- round(data, digits = 4)
summary(datavis)
```



#EXPLORATORY DATA ANALYSIS

#Frequency Distribution of Quality
```{r}
#frequency plot for quality

ggplot(data, aes(factor(quality))) + geom_bar(fill = "light blue") + labs(x = "Quality", y="Frequency", title = "Frequency Distribution for Quality Ratings") + theme( plot.title=element_text(hjust=0.5))
```


#Correlation Matrix
```{r}
#correlation matrix 

ggcorrplot(cor(data), lab=FALSE)
```


#Highly Correlated Regressors
```{r}
#investigating the higher correlations between regressors in the correlation matrix to avoid multicollinearity
#define highly correlated as cor >= 0.3

multicolcor <- abs(cor(data)) >= 0.3

freeVtot <- cor(data$free.sulfur.dioxide, data$total.sulfur.dioxide)
denVfa <- cor(data$density, data$fixed.acidity)
caVfa <- cor(data$citric.acid, data$fixed.acidity)
pHVden <- cor(data$pH, data$density)
pHVca <- cor(data$pH, data$citric.acid)
alcVden <- cor(data$alcohol, data$density)
caVva <- cor(data$citric.acid, data$volatile.acidity)
pHVfa <- cor(data$pH, data$fixed.acidity)
caVden <- cor(data$citric.acid, data$density)
caVsulph <- cor(data$citric.acid, data$sulphates)
rsVden <- cor(data$residual.sugar, data$density)
chlorVsulph <- cor(data$chlorides, data$sulphates)

cors <- c(caVfa, denVfa, freeVtot, chlorVsulph, caVden, rsVden, caVsulph, pHVden, alcVden, pHVca, caVva, pHVfa)

cornames <- c("ca v fa", "den v fa", "free v tot", "chlor v sulph", "ca v den", "rs v den", "ca v sulph", "pH vs den", "alc v den", "pH v ca", "ca v va", "pH v fa")

names(cors) <- cornames

cors
```


#Correlation of each dependent variable with Quality
```{r}
#basic correlation chart with quality

corfa <- cor(data$quality, data$fixed.acidity)
corva <- cor(data$quality, data$volatile.acidity)
corca <- cor(data$quality, data$citric.acid)
corrs <- cor(data$quality, data$residual.sugar)
corchlor <- cor(data$quality, data$chlorides)
corfree <- cor(data$quality, data$free.sulfur.dioxide)
cortot <- cor(data$quality, data$total.sulfur.dioxide)
corden <- cor(data$quality, data$density)
corpH <- cor(data$quality, data$pH)
corsulph <- cor(data$quality, data$sulphates)
coralc <- cor(data$quality, data$alcohol)

qualitycors <- c(coralc, corsulph, corca, corfa, corrs, corfree, corpH, corchlor, corden, cortot, corva)

qualitycornames <- c("alc", "sulph", "ca", "fa", "rs", "free", "pH", "chlor", "den", "tot", "va")

names(qualitycors) <- qualitycornames

qualitycors
```


#Boxplots of each dependent varaible with Quality
```{r}
#relationship between sugar content and quality

ggplot(datavis, aes(x=factor(quality), y=residual.sugar, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() + 
  labs(x = "Quality", y= "Residual Sugar", title="Sugar Content vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()
```


```{r}
#relationship between alcohol content and quality

ggplot(datavis, aes(x=factor(quality), y=alcohol, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() + 
  labs(x = "Quality", y= "Alcohol Content", title="Alcohol vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()
```


```{r}
#relationship between chlorides and quality

ggplot(datavis, aes(x=factor(quality), y=chlorides, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() + 
  labs(x = "Quality", y= "Chlorides", title="Chlorides vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()
```


```{r}
#relationship between density and quality

ggplot(datavis, aes(x=factor(quality), y=density, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() + 
  labs(x = "Quality", y= "Density", title="Density vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()
```


```{r}
#relationship between pH and quality

ggplot(datavis, aes(x=factor(quality), y=pH, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() + 
  labs(x = "Quality", y= "pH", title="pH vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()
```


```{r}
#relationship between sulphates and quality

ggplot(datavis, aes(x=factor(quality), y=sulphates, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() + 
  labs(x = "Quality", y= "Sulphates", title="Sulphates vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()

```


```{r}
#relationship between sulfur dioxide content and quality

total <- ggplot(datavis, aes(x=factor(quality), y=total.sulfur.dioxide, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "Quality", y= "Total Sulfur Dioxide", title="Total Sulfur Dioxide vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()

free <- ggplot(datavis, aes(x=factor(quality), y=free.sulfur.dioxide, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "Quality", y= "Free Sulfur Dioxide", title="Free Sulfur Dioxide vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()

ggarrange(free, total, row = 1, col = 2)
```


```{r}
#relationship between acidic content and quality

fixed <- ggplot(datavis, aes(x=factor(quality), y=fixed.acidity, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "Quality", y= "Fixed Acidity", title="Fixed Acidity vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()

volatile <- ggplot(datavis, aes(x=factor(quality), y=volatile.acidity, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "Quality", y= "Volatile Acidity", title="Volatile Acidity vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()

citric <- ggplot(datavis, aes(x=factor(quality), y=citric.acid, group=factor(quality), fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "Quality", y= "Citric Acid", title="Citric Acid vs. Quality") + 
  theme(legend.position="none", 
        plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) + coord_flip()

ggarrange(fixed, volatile, citric, row = 1, col = 3)
```


#Categorizing wines as high-quality, medium-quality, or low-quality
```{r}
#creating a new column for categorizing wines by high, medium, or low quality ratings

qual <- data$quality
qualitytype <- c()

for (x in qual) {
  if (x <= 4) {qualitytype = c(qualitytype, "low")}
  else if(x >= 7) {qualitytype = c(qualitytype, "high")}
  else {qualitytype = c(qualitytype, "medium")}
}

qualityfactor <- factor(qualitytype, ordered = TRUE, levels = c("low", "medium", "high"))
data1 <- cbind(data, qualityfactor)
summary(data1)
```


```{r}
#frequency plot for quality types

ggplot(data1, aes(factor(qualitytype))) + geom_bar(fill = "blue") + labs(x = "Quality Type", y="Frequency", title = "Frequency Distribution for Quality Types") + theme( plot.title=element_text(hjust=0.5))
```


#Density Plots of each dependent variable by Quality Type
```{r}
#density plot for pH compared to quality

ggplot(data1, aes(x= pH, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of pH by Quality Type"))
```


```{r}
#density plot for density of wine compared to quality

ggplot(data1, aes(x= density, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of Wine by Quality Type"))
```


```{r}
#density plot for alcohol content compared to quality

ggplot(data1, aes(x= alcohol, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of Alcohol Content by Quality Type"))
```


```{r}
#density plot for sulphates compared to quality

ggplot(data1, aes(x=sulphates, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of Sulphate Content by Quality Type"))
```


```{r}
#density plot for fixed acidity compared to quality

ggplot(data1, aes(x= fixed.acidity, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of Fixed Acidity by Quality Type"))
```


```{r}
#density plot for volatile acidity compared to quality

ggplot(data1, aes(x= volatile.acidity, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of Volatile Acidity by Quality Type"))
```


```{r}
#density plot for citric acid compared to quality

ggplot(data1, aes(x=citric.acid, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of Citric Acid by Quality Type"))
```


```{r}
#density plot for residual sugar compared to quality

ggplot(data1, aes(x= residual.sugar, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of Residual Sugar by Quality Type"))
```


```{r}
#density plot for chlorides compared to quality

ggplot(data1, aes(x= chlorides, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of Chlorides by Quality Type"))
```


```{r}
#density plot for free sulfur dioxide compared to quality

ggplot(data1, aes(x= free.sulfur.dioxide, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of Free Sulfur Dioxide by Quality Type"))
```


```{r}
#density plot for total sulfur dioxide compared to quality

ggplot(data1, aes(x= total.sulfur.dioxide, fill = qualityfactor)) + 
  geom_density(alpha = 0.5) + 
  theme(legend.title = element_blank(), plot.title = element_text(hjust=0.5)) + 
  scale_fill_discrete(labels = c("low", "medium","high"))  + 
  labs(title = element_text("Density of Total Sulfur Dioxide by Quality Type"))
```


#Scatterplots of Quality of Individual Wines by each dependent variable
```{r}
#printing out one instance of each graph with legend for the corresponding scatterplots 

#the remainder of the scatterplots will not have a legend for the sake of space, but one can see the legend here, and it is the same for all of the following scatterplots 

legendfactors <- ggplot(data = data1, aes(ID, alcohol)) + geom_point(aes(color = qualityfactor), show.legend = TRUE) + labs(x= "Wines", y = "Alcohol") + theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank())

legendcontinuous <- ggplot(data = data, aes(ID, alcohol)) + geom_point(aes(color = quality), show.legend = TRUE) + labs(x= "Wines", y = "Alcohol Content") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

#graph with legend for factored wine quality
legendfactors
```


```{r}
#graph with legend for continuous wine quality
legendcontinuous
```


```{r}
#scatterplot of alcohol content of individual wines by quality ratings (factored)

alcf <- ggplot(data = data1, aes(ID, alcohol)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "Alcohol") + theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank())

#scatterplot of alcohol content of individual wines by quality ratings (continuous)

alcc <- ggplot(data = data, aes(ID, alcohol)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y = "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(alcf, alcc, row = 1, col = 2)
```


```{r}
#scatterplot of sulphates of individual wines by quality ratings (factored)

sulphf <- ggplot(data = data1, aes(ID, sulphates)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "Sulphates") + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

#scatterplot of sulphates of individual wines by quality ratings (continuous)

sulphc <- ggplot(data = data, aes(ID, sulphates)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y = "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(sulphf, sulphc, row = 1, col = 2)
```


```{r}
#scatterplot of pH of individual wines by quality ratings (factored)

pHf <- ggplot(data = data1, aes(ID, pH)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "pH") + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

#scatterplot of pH of individual wines by quality ratings (continuous)

pHc <- ggplot(data = data, aes(ID, pH)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y = "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(pHf, pHc, row = 1, col = 2)
```


```{r}
#scatterplot of density of individual wines by quality ratings (factored)

densityf <- ggplot(data = data1, aes(ID, density)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "Density") + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

#scatterplot of density of individual wines by quality ratings (continuous)

densityc <- ggplot(data = data, aes(ID, density)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y = "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(densityf, densityc, row = 1, col = 2)
```


```{r}
#scatterplot of fixed acidity of individual wines by quality ratings (factored)

faf <- ggplot(data = data1, aes(ID, fixed.acidity)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "Fixed Acidity") + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

#scatterplot of fixed acidity of individual wines by quality ratings (continuous)

fac <- ggplot(data = data, aes(ID, fixed.acidity)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y = "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(faf, fac, row = 1, col = 2)
```


```{r}
#scatterplot of volatile acidity of individual wines by quality ratings (factored)

vaf <- ggplot(data = data1, aes(ID, volatile.acidity)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "Volatile Acidity") + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

#scatterplot of volatile acidity of individual wines by quality ratings (continuous)

vac <- ggplot(data = data, aes(ID, volatile.acidity)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y = "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(vaf, vac, row = 1, col = 2)
```


```{r}
#scatterplot of citric acid of individual wines by quality ratings (factored)

caf <- ggplot(data = data1, aes(ID, citric.acid)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "Citric Acid") + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

#scatterplot of citric acid of individual wines by quality ratings (continuous)

cac <- ggplot(data = data, aes(ID, citric.acid)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y = "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(caf, cac, row = 1, col = 2)
```


```{r}
#scatterplot of residual sugar of individual wines by quality ratings (factored)

rsf <- ggplot(data = data1, aes(ID, residual.sugar)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "Residual Sugar") + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

#scatterplot of residual sugar of individual wines by quality ratings (continuous)

rsc <- ggplot(data = data, aes(ID, residual.sugar)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y = "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(rsf, rsc, row = 1, col = 2)
```


```{r}
#scatterplot of chlorides of individual wines by quality ratings (factored)

chlorf <- ggplot(data = data1, aes(ID, chlorides)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "Chlorides") + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

#scatterplot of chlorides of individual wines by quality ratings (continuous)

chlorc <- ggplot(data = data, aes(ID, chlorides)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y = "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(chlorf, chlorc, row = 1, col = 2)
```


```{r}
#scatterplot of free sulfur dioxide of individual wines by quality ratings (factored)

freesulff <- ggplot(data = data1, aes(ID, free.sulfur.dioxide)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "Free Sulfur Dioxide") + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

#scatterplot of free sulfur dioxide of individual wines by quality ratings (continuous)

freesulfc <- ggplot(data = data, aes(ID, free.sulfur.dioxide)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y = "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(freesulff, freesulfc, row = 1, col = 2)
```


```{r}
#scatterplot of total sulfur dioxide of individual wines by quality ratings (factored)

totsulff <- ggplot(data = data1, aes(ID, total.sulfur.dioxide)) + geom_point(aes(color = qualityfactor), show.legend = FALSE) + labs(x= "Wines", y = "Total Sulfur Dioxide") + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

#scatterplot of total sulfur dioxide of individual wines by quality ratings (continuous)

totsulfc <- ggplot(data = data, aes(ID, total.sulfur.dioxide)) + geom_point(aes(color = quality), show.legend = FALSE) + labs(x= "Wines", y= "") + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

ggarrange(totsulff, totsulfc, row = 1, col = 2)
```



#ANALYSIS

#Bootstrapping the data
```{r}
#setting a seed so that the analysis is reproducible
set.seed(1998)  

#setting up a train/test dataset 
##using convention of 80% train data, 20% test data
ndata <- nrow(data)
train_index <- sample(1:ndata, size=0.8*ndata, rep=FALSE)
test_index <- setdiff(1:ndata, train_index)
train <- data[train_index,]
test <- data[test_index,]

#bootstrapping 
set.seed(1899)
trainboot <- train[1,]
trainboot[,] <- NA
for (i in 1:3200) {
  trainboot[i,] <- train[sample(1:1279, replace = TRUE,1),]
}

str(trainboot)
```



#MODELING

#Linear Regression Models
```{r}
#model1 with all regressors included, original data

lm1 <- lm(quality ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = data)

summary(lm1)
```


```{r}
#model2 with only significant regressors included, original data 
#significant regessors are: volatile acidity, chlorides, free sulfur dioxide, total sulfur dioxide, pH, sulphates, alcohol

lm2 <- lm(quality ~ volatile.acidity + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + pH + sulphates + alcohol, data = data)

summary(lm2)
```


```{r}
#model3 with all regressors included, bootstrap data

lm3 <- lm(quality ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = trainboot)

summary(lm3)
```


```{r}
#model4 with only significant regressors included, bootstrap data
#significant regressors are: fixed acidity, volatile acidity, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, sulphates, alcohol

lm4 <- lm(quality ~ fixed.acidity + volatile.acidity + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + sulphates + alcohol, data = trainboot)

summary(lm4)
```


#Random Forest Models
```{r}
#running a random forest with the original data

model5 <- randomForest(quality ~ volatile.acidity + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + pH + sulphates + alcohol, data = data, nodesize = 3, ntree = 1000, mtry = 7)

model5
```


```{r}
#running a random forest with the bootstrapped data

model6 <- randomForest(quality ~ volatile.acidity + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + pH + sulphates + alcohol, data = trainboot, nodesize = 3, ntree = 1000, mtry = 7)

model6 
```


#Classification Trees
```{r}
#classification tree on original data

model7 <- tree(quality ~ volatile.acidity + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + pH + sulphates + alcohol, data = data)
par(xpd = NA) 
plot(model7) + text(model7)
```


```{r}
#classification tree on bootstrapped data

model8 <- tree(quality ~ volatile.acidity + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + pH + sulphates + alcohol, data = trainboot)
par(xpd = NA) 
plot(model8) + text(model8)
```


```{r}
#CART on original data

model9 <- rpart(quality ~ volatile.acidity + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + pH + sulphates + alcohol, data = data)
par(xpd = NA) 
plot(model9) + text(model9)
```


```{r}
#CART on bootstrapped data

model10 <- rpart(quality ~ volatile.acidity + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + pH + sulphates + alcohol, data = trainboot)
par(xpd = NA) 
plot(model10) + text(model10)

```



#EVALUATION

#evaluations only performed with models 1-6 since the trees for models 7-10 are not evaluated similarly and thus cannot be compared

```{r}
#running predictions on the linear models using original data
predictions1 <- predict(lm1)
predictions2 <- predict(lm2)

#running predictions of the bootstrap models on the test set
predictions3 <- predict(lm3, newdata=test)
predictions4 <- predict(lm4, newdata=test)
true_values <- test$quality

#running predictions of the randomForest models
predictions5 <- predict(model5)
predictions6 <- predict(model6, newdata=test)
```


```{r}
#mean absolute error
MAE <- function(prediction, true_values) {
    return (mean(abs(prediction-true_values)))
}

MAE1 <- MAE(predictions1, true_values)
MAE2 <- MAE(predictions2, true_values)
MAE3 <- MAE(predictions3, true_values)
MAE4 <- MAE(predictions4, true_values)
MAE5 <- MAE(predictions5, true_values)
MAE6 <- MAE(predictions6, true_values)

allMAE <- c(round(MAE1, digits = 4), round(MAE2, digits = 4), round(MAE3, digits = 4), round(MAE4, digits = 4), round(MAE5, digits = 4), round(MAE6, digits = 4))
namedMAE <- c("model1", "model2", "model3", "model4", "model5", "model6")
names(allMAE) <- namedMAE
allMAE
```


```{r}
#root mean squared error
RMSE <- function(predictions, true_values) {
    return (sqrt(mean((predictions-true_values)^2)))
}

RMSE1 <- RMSE(predictions1, true_values)
RMSE2 <- RMSE(predictions2, true_values)
RMSE3 <- RMSE(predictions3, true_values)
RMSE4 <- RMSE(predictions4, true_values)
RMSE5 <- RMSE(predictions5, true_values)
RMSE6 <- RMSE(predictions6, true_values)

allRMSE <- c(round(RMSE1, digits = 4), round(RMSE2, digits = 4), round(RMSE3, digits = 4), round(RMSE4, digits = 4), round(RMSE5, digits = 4), round(RMSE6, digits = 4))
namedRMSE <- c("model1", "model2", "model3", "model4", "model5", "model6")
names(allRMSE) <- namedRMSE
allRMSE

```


```{r}
#mean absolute percentage error
MAPE <- function(predictions, true_values) {
    return (mean(abs((predictions - true_values) / true_values)))
}

MAPE1 <- MAPE(predictions1, true_values)
MAPE2 <- MAPE(predictions2, true_values)
MAPE3 <- MAPE(predictions3, true_values)
MAPE4 <- MAPE(predictions4, true_values)
MAPE5 <- MAPE(predictions5, true_values)
MAPE6 <- MAPE(predictions6, true_values)

allMAPE <- c(round(MAPE1, digits = 4), round(MAPE2, digits = 4), round(MAPE3, digits = 4), round(MAPE4, digits = 4), round(MAPE5, digits = 4), round(MAPE6, digits = 4))
namedMAPE <- c("model1", "model2", "model3", "model4", "model5", "model6")
names(allMAPE) <- namedMAPE
allMAPE

```


```{r}
#AIC scores for linear models 1-4
AIC1 <- AIC(lm1)
AIC2 <- AIC(lm2)
AIC3 <- AIC(lm3)
AIC4 <- AIC(lm4)

allAIC <- c(round(AIC1, digits = 4), round(AIC2, digits = 4), round(AIC3, digits = 4), round(AIC4, digits = 4))
namedAIC <- c("model1", "model2", "model3", "model4")
names(allAIC) <- namedAIC
allAIC
```


```{r}
#exploration of potential multicollinearity in models
#based on EDA, there may be a multicollinearity issue with some of the regressors

#VIF scores
VIF1 <- vif(lm1)
VIF2 <- vif(lm2)
VIF3 <- vif(lm3)
VIF4 <- vif(lm4)

VIF1
VIF2
VIF3
VIF4
#no VIF scores > 10
```


